---
title: "RTC Fact Sheet: 2025 State of Evictions"
format: 
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false

library(DBI)
library(RPostgres)
library(flextable)
library(sf)
library(tidyverse)
library(socratadata)
library(lubridate)
library(writexl)

# Ensure that the Quarto render and RStudio read the same environment vars
readRenviron(".Renviron")

con <- dbConnect(
  RPostgres::Postgres(),
  dbname = Sys.getenv("PG_DBNAME"),
  host     = Sys.getenv("PG_HOST"),
  port     = Sys.getenv("PG_PORT"),
  user     = Sys.getenv("PG_USER"),
  password = Sys.getenv("PG_PASSWORD")
  )
```


The following analyses were performed in support of [Right to Counsel's](https://www.righttocounselnyc.org/) annual State of Evictions fact sheet. 

```{r, cache=TRUE, message=FALSE}
# First, pull data. Then, work locally in R.

### Evictions db SQL queries ###

# Eviction Filings, 2020 - 2025, all
filings_20_25 <- dbGetQuery(con, "
  SELECT DISTINCT(i.indexnumberid), fileddate, court
  FROM oca_index i
  WHERE i.fileddate >= MAKE_DATE(2020,1,1) 
      AND i.fileddate <= MAKE_DATE(2025,12,31)
      AND i.classification IN ('Holdover', 'Non-Payment')
      AND (i.propertytype = 'Residential' OR i.propertytype IS NULL);
")

# Eviction Filings, 2020 - 2025, with location
filings_20_25_sf <- st_read(con, query = "
  SELECT DISTINCT(i.indexnumberid), fileddate, court,
    ST_Transform(a.geom, 2263) AS geom
  FROM oca_index i
  JOIN oca_addresses a ON i.indexnumberid = a.indexnumberid
  WHERE i.fileddate >= MAKE_DATE(2020,1,1) 
      AND i.fileddate <= MAKE_DATE(2025,12,31)
      AND i.classification IN ('Holdover', 'Non-Payment')
      AND (i.propertytype = 'Residential' OR i.propertytype IS NULL)
      AND a.geom IS NOT NULL;
")

# Executed Evictions, 2020 - 2025, all
executions_20_25 <- dbGetQuery(con, "
  SELECT DISTINCT(i.indexnumberid), w.executiondate, i.court
  FROM oca_index i
  JOIN oca_warrants w ON i.indexnumberid = w.indexnumberid
  WHERE w.executiondate >= MAKE_DATE(2020,1,1) 
    AND w.executiondate <= MAKE_DATE(2025,12,31)
    AND i.classification IN ('Holdover', 'Non-Payment')
    AND (i.propertytype = 'Residential' OR i.propertytype IS NULL)
    AND w.executiondate IS NOT NULL;
")

# Executed Evictions, 2020 - 2025, with location
executions_20_25_sf <- st_read(con, query = "
  WITH executed_evictions AS (
  	SELECT DISTINCT(i.indexnumberid), w.executiondate, i.court
  	FROM oca_index i
  	JOIN oca_warrants w ON i.indexnumberid = w.indexnumberid
  	WHERE w.executiondate >= MAKE_DATE(2020,1,1) 
  	  AND w.executiondate <= MAKE_DATE(2025,12,31)
  	  AND i.classification IN ('Holdover', 'Non-Payment')
  	  AND (i.propertytype = 'Residential' OR i.propertytype IS NULL)
  	  AND w.executiondate IS NOT NULL
  )
  SELECT e.*, ST_TRANSFORM(a.geom, 2263)
  FROM executed_evictions e
  JOIN oca_addresses a on e.indexnumberid = a.indexnumberid
  WHERE a.geom IS NOT NULL
")

### Boundary Files ###
# NYC Boundary via borough boundary union
nyc <- read_sf("https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_Borough_Boundary/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson") |> 
  st_transform(2263) |> 
  st_make_valid() |> 
  st_union()

# NYS Assembly
nysad <- read_sf("https://services6.arcgis.com/EbVsqZ18sv1kVJ3k/ArcGIS/rest/services/NYS_Assembly_Districts/FeatureServer/0/query?where=0%3D0&outFields=*&f=geojson") |> 
  st_transform(2263) |> 
  st_make_valid()

# NYS Senate
nyss <- read_sf("https://services6.arcgis.com/EbVsqZ18sv1kVJ3k/ArcGIS/rest/services/NYS_Senate_Districts/FeatureServer/0/query?where=0%3D0&outFields=*&f=geojson") |> 
  st_transform(2263) |> 
  st_make_valid()

# NYC Council
nyc_council <- read_sf("https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_City_Council_Districts/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson") |> 
  st_transform(2263) |> 
  st_make_valid()

# NYC Council members
nyc_council_members <- soc_read(
  "https://data.cityofnewyork.us/City-Government/City-Council-Members-1999-to-Present-/uvw5-9znb.json"
) |> 
  # Filter to current term
  filter(Sys.Date() %within% interval(start = term_start, end = term_end))

# Join CMs to spatial file
nyc_council <- nyc_council |> 
  left_join(nyc_council_members |> 
              select(name, district),
            by = join_by(CounDist == district)) |> 
  # Designate Vacant seats
  mutate(name = replace_na(name, "Vacant"))

# NYC Boundaries via Borough union
bb <- read_sf("https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_Borough_Boundary/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson") |> 
  st_transform(2263) |> 
  st_make_valid() |> 
  st_union()

# NYS Counties
nys_counties <- read_sf("https://services6.arcgis.com/EbVsqZ18sv1kVJ3k/ArcGIS/rest/services/NYS_Civil_Boundaries/FeatureServer/2/query?where=0%3D0&outFields=*&f=geojson") |> 
  st_transform(2263)

# Separate out upstate / downstate
nysad_downstate <- st_filter(nysad, bb, .predicate = st_intersects)

nysad_upstate <- st_filter(nysad, bb, .predicate = st_disjoint)

nyss_downstate <- st_filter(nyss, bb, .predicate = st_intersects)
  
nyss_upstate <- st_filter(nyss, bb, .predicate = st_disjoint)
```


## Cumulative Eviction Filings

### Number of eviction filings from January 1, 2020 - December 31, 2025

In New York State, there were **`r scales::comma(nrow(filings_20_25))`** eviction-oriented (holdover and non-payment) filings from January 1, 2020 to December 31, 2025.

Please note that this analysis **excludes commercial evictions within New York City.** Unfortunately, outside of NYC, excluding commercial from residential eviction filings and executions is not possible given the current data.

## Annual eviction filings
```{r}
eviction_count_annual_df <- filings_20_25 |> 
  group_by(filed_year = year(fileddate)) |> 
  summarize(count = n())
```

The number of eviction-related (holdover and non-payment) eviction filings per year is:

```{r}
eviction_count_annual_table <- eviction_count_annual_df |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    filed_year = "Year",
    count = "Eviction Filings"
  ) |> 
  colformat_num(j=1,
                big.mark = "") |> 
  align(j=1,
        align = "left",
        part = "all")

eviction_count_annual_table
```

## Which areas are hardest hit?

### By court, January 1, 2020 - December 31, 2025:
```{r}
eviction_count_courts_df <- filings_20_25 |> 
  group_by(court) |> 
  summarize(count = n()) |> 
  arrange(desc(count))

eviction_count_courts_table <- eviction_count_courts_df |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    court = "Court",
    count = "Eviction Filings"
  )

eviction_count_courts_table
```

### By court, January 1, 2025 - December 31, 2025:
```{r}
eviction_count_courts_25_df <- filings_20_25 |> 
  filter(year(fileddate) == 2025) |> 
  group_by(court) |> 
  summarize(count = n()) |> 
  arrange(desc(count))

eviction_count_courts_25_table <- eviction_count_courts_25_df |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    court = "Court",
    count = "Eviction Filings"
  )

eviction_count_courts_25_table
```

### By NYS Assembly District: January 1, 2020 - December 31, 2025

*In the following filing counts by area, note that only `r scales::comma(nrow(filings_20_25_sf))` eviction filings (`r round((nrow(filings_20_25_sf)/nrow(filings_20_25)*100),2)`%) have a location*

#### Upstate Assembly Districts (Not intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nysad_upstate_5year <- filings_20_25_sf |> 
  st_join(nysad_upstate |> 
              select(District, Name),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(District, Name) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nysad_upstate_5year_table <- nysad_upstate_5year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    Name = "Representative"
  )

nysad_upstate_5year_table
```

#### Downstate Assembly Districts (Intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nysad_downstate_5year <- filings_20_25_sf |> 
  st_join(nysad_downstate |> 
              select(District, Name),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(District, Name) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nysad_downstate_5year_table <- nysad_downstate_5year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all")

nysad_downstate_5year_table
```

### By NYS Assembly District: January 1, 2025 - December 31, 2025

#### Upstate Assembly Districts (Not intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nysad_upstate_1year <- filings_20_25_sf |> 
  filter(year(fileddate) == 2025) |> 
  st_join(nysad_upstate |> 
              select(District, Name),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(District, Name) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nysad_upstate_1year_table <- nysad_upstate_1year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all")

nysad_upstate_1year_table
```

#### Downstate Assembly Districts (Intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nysad_downstate_1year <- filings_20_25_sf |> 
  filter(year(fileddate) == 2025) |>  
  st_join(nysad_downstate |> 
              select(District, Name),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(District, Name) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`)) 

nysad_downstate_1year_table <- nysad_downstate_1year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all")

nysad_downstate_1year_table
```

### By NYS Senate District, January 1, 2020 - December 31, 2025

#### Upstate Senate Districts (Not intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nyss_upstate_5year <- filings_20_25_sf |> 
  st_join(nyss_upstate |> 
              select(DISTRICT, NAME),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(DISTRICT, NAME) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`)) 

nyss_upstate_5year_table <- nyss_upstate_5year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    DISTRICT = "District",
    NAME = "Representative"
  )

nyss_upstate_5year_table
```

#### Downstate Senate Districts (Intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nyss_downstate_5year <- filings_20_25_sf |> 
  st_join(nyss_downstate |> 
              select(DISTRICT, NAME),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(DISTRICT, NAME) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`)) 

nyss_downstate_5year_table <- nyss_downstate_5year |>
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    DISTRICT = "District",
    NAME = "Representative"
  )

nyss_downstate_5year_table
```

### By NYS Senate District, January 1, 2025 - December 31, 2025
#### Upstate Senate Districts (Not intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nyss_upstate_1year <- filings_20_25_sf |> 
  filter(year(fileddate) == 2025) |> 
  st_join(nyss_upstate |> 
              select(DISTRICT, NAME),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(DISTRICT, NAME) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nyss_upstate_1year_table <- nyss_upstate_1year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    DISTRICT = "District",
    NAME = "Representative"
  )

nyss_upstate_1year_table
```

#### Downstate Senate Districts (Intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nyss_downstate_1year <- filings_20_25_sf |> 
  filter(year(fileddate) == 2025) |> 
  st_join(nyss_downstate |> 
              select(DISTRICT, NAME),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(DISTRICT, NAME) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nyss_downstate_1year_table <- nyss_downstate_1year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    DISTRICT = "District",
    NAME = "Representative"
  )

nyss_downstate_1year_table
```

### Executed Evictions in 2025

Note: Of the `r scales::comma(nrow(executions_20_25 |> filter(year(executiondate)==2025)))` evictions executed statewide in 2025, only `r scales::comma(nrow(executions_20_25_sf |> filter(year(executiondate)==2025)))` have a location associated with them (`r round((nrow(executions_20_25_sf |> filter(year(executiondate)==2025))/nrow(executions_20_25 |> filter(year(executiondate)==2025))*100),2)`%)

``` {r, cache=FALSE}
executed_evictions_25_byCty <- executions_20_25_sf |> 
  filter(year(executiondate) == 2025) |> 
  st_join(nys_counties |> 
            select(NAME)) |> 
  st_drop_geometry() |> 
  group_by(NAME) |> 
  summarize(`Executed Evictions` = n()) |> 
  arrange(desc(`Executed Evictions`))

executed_evictions_25_byCty_table <- executed_evictions_25_byCty |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    NAME = "County"
  )

executed_evictions_25_byCty_table
```


### Evictions During the Winter Months (November 1 - April 15)

#### Eviction filings

```{r}
filings_20_25 <- filings_20_25 |> 
  mutate(winter_season = case_when(month(fileddate) >= 11 ~ TRUE,
                                   month(fileddate) < 4 ~ TRUE,
                                   month(fileddate) == 4 & day(fileddate) <= 15 ~ TRUE,
                                   .default = FALSE
                                   ))

winter_filings_table <- filings_20_25 |> 
  group_by(filed_year = year(fileddate)) |> 
  summarize(all_filings = n(),
            winter_filings = sum(winter_season)) |> 
  mutate(winter_filing_pct = round((winter_filings/all_filings*100), 1)) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    filed_year = "Year",
    all_filings = "Annual Filings",
    winter_filings = "Winter Filings",
    winter_filing_pct = "Winter Filing Percentage (%)"
  ) |> 
  colformat_num(j=1,
                big.mark = "") |> 
  colformat_int(j=2:3,
                big.mark = ",") |> 
  align(j=1,
      align = "left",
      part = "all")

winter_filings_table
```

#### Executed evictions

```{r}
executions_20_25 <- executions_20_25 |> 
  mutate(winter_season = case_when(month(executiondate) >= 11 ~ TRUE,
                                   month(executiondate) < 4 ~ TRUE,
                                   month(executiondate) == 4 & day(executiondate) <= 15 ~ TRUE,
                                   .default = FALSE
                                   ))

winter_executions_table <- executions_20_25 |> 
  group_by(executed_year = year(executiondate)) |>
  summarize(all_executions = n(),
            winter_executions = sum(winter_season)) |> 
  mutate(winter_execution_pct = round(winter_executions/all_executions*100,1)) |> 
  flextable() |>
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    executed_year = "Year",
    all_executions = "Annual Executions",
    winter_executions = "Winter Executions",
    winter_execution_pct = "Winter Execution Percentage (%)"
  ) |> 
  colformat_num(j=1,
                big.mark = "") |> 
  colformat_int(j=2:3,
                big.mark = ",") |> 
  align(j=1,
      align = "left",
      part = "all")
  
winter_executions_table 
```

Compiling filings, executions, winter filings, and winter executions for the two political districts plus NYC Council districts:
```{r}
# Compile filings, executions, winter filings, winter executions for all political districts
filings_20_25_sf <- filings_20_25_sf |> 
  mutate(winter_season = case_when(month(fileddate) >= 11 ~ TRUE,
                                   month(fileddate) < 4 ~ TRUE,
                                   month(fileddate) == 4 & day(fileddate) <= 15 ~ TRUE,
                                   .default = FALSE
                                   ))

executions_20_25_sf <- executions_20_25_sf |> 
  mutate(winter_season = case_when(month(executiondate) >= 11 ~ TRUE,
                                   month(executiondate) < 4 ~ TRUE,
                                   month(executiondate) == 4 & day(executiondate) <= 15 ~ TRUE,
                                   .default = FALSE
                                   ))

### Assembly Districts ###
nysad_compiled <- nysad |> 
  select(District, Name, Party) |> 
  mutate(Region = case_when(District %in% nysad_upstate$District ~ "Upstate",
                            District %in% nysad_downstate$District ~ "Downstate",
                            .default = NA))

executions_20_25_nysad <- executions_20_25_sf |> 
  st_join(nysad_compiled |> select(District))



nysad_compiled <- nysad_compiled |> 
  left_join(
    bind_rows(
      nysad_downstate_5year,
      nysad_upstate_5year
    ) |> 
      rename("filings_20_25" = "Eviction Filings") |> 
      select(!Name),
    by = join_by(District)
  ) |> 
  left_join(
    bind_rows(
      nysad_downstate_1year,
      nysad_upstate_1year
    ) |> 
      rename("filings_25" = "Eviction Filings") |> 
      select(!Name),
    by = join_by(District)
  ) |> 
  left_join(executions_20_25_nysad |> 
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(executions_20_25 = n())) |> 
  left_join(executions_20_25_nysad |> 
              filter(year(executiondate) == 2025) |> 
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(executions_25 = n())) |> 
  left_join(filings_20_25_sf |> 
              filter(winter_season == TRUE) |> 
              st_join(nysad_compiled |> select(District)) |>
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(winter_filings_20_25 = n())) |> 
  left_join(filings_20_25_sf |> 
              filter(winter_season == TRUE & year(fileddate) == 2025) |> 
              st_join(nysad_compiled |> select(District)) |>
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(winter_filings_25 = n())) |> 
  left_join(executions_20_25_nysad |> 
              filter(winter_season == TRUE) |> 
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(winter_executions_20_25 = n())) |> 
  left_join(executions_20_25_nysad |> 
              filter(winter_season == TRUE & year(executiondate) == 2025) |> 
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(winter_executions_25 = n())) |> 
  st_drop_geometry()

# Rearrange
nysad_compiled <- nysad_compiled |> 
  relocate(contains("20_25"),
           .after = Region)

### Senate Districts ###
nyss_compiled <- nyss |> 
  select(DISTRICT, NAME, Party) |> 
  mutate(Region = case_when(DISTRICT %in% nyss_upstate$DISTRICT ~ "Upstate",
                            DISTRICT %in% nyss_downstate$DISTRICT ~ "Downstate",
                            .default = NA))

executions_20_25_nyss <- executions_20_25_sf |> 
  st_join(nyss_compiled |> select(DISTRICT))

nyss_compiled <- nyss_compiled |> 
  left_join(
    bind_rows(
      nyss_upstate_5year,
      nyss_downstate_5year
    ) |> 
      rename("filings_20_25" = "Eviction Filings") |> 
      select(!NAME),
    by = join_by(DISTRICT)
  ) |> 
  left_join(
    bind_rows(
      nyss_downstate_1year,
      nyss_upstate_1year
    ) |> 
      rename("filings_25" = "Eviction Filings") |> 
      select(!NAME),
    by = join_by(DISTRICT)
  ) |> 
  left_join(executions_20_25_nyss |> 
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(executions_20_25 = n())) |> 
  left_join(executions_20_25_nyss |> 
              filter(year(executiondate) == 2025) |> 
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(executions_25 = n())) |> 
  left_join(filings_20_25_sf |> 
              filter(winter_season == TRUE) |> 
              st_join(nyss_compiled |> select(DISTRICT)) |>
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(winter_filings_20_25 = n())) |> 
  left_join(filings_20_25_sf |> 
              filter(winter_season == TRUE & year(fileddate) == 2025) |> 
              st_join(nyss_compiled |> select(DISTRICT)) |>
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(winter_filings_25 = n())) |> 
  left_join(executions_20_25_nyss |> 
              filter(winter_season == TRUE) |> 
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(winter_executions_20_25 = n())) |> 
  left_join(executions_20_25_nyss |> 
              filter(winter_season == TRUE & year(executiondate) == 2025) |> 
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(winter_executions_25 = n())) |> 
  st_drop_geometry()
  
# Rearrange
nyss_compiled <- nyss_compiled |> 
  relocate(contains("20_25"),
           .after = Region)


### NYC Council ###
nyc_council <- nyc_council |> 
  select(CounDist, name) |> 
  arrange(CounDist)

filings_20_25_nycc <- filings_20_25_sf |> 
  st_join(nyc_council |> select(CounDist)) |> 
  st_drop_geometry()

executions_20_25_nycc <- executions_20_25_sf |> 
  st_join(nyc_council |> select(CounDist)) |> 
  st_drop_geometry()

nyc_council_compiled <- nyc_council |> 
  # Filings 20-25
  left_join(filings_20_25_nycc |> 
              filter(!is.na(CounDist)) |> 
              group_by(CounDist) |> 
              summarize(filings_20_25 = n())) |> 
  # Executions 20-25
  left_join(executions_20_25_nycc |> 
              filter(!is.na(CounDist)) |> 
              group_by(CounDist) |> 
              summarize(executions_20_25 = n())) |> 
  # Winter filings 20-25
  left_join(filings_20_25_nycc |> 
              filter(!is.na(CounDist) & winter_season == TRUE) |> 
              group_by(CounDist) |> 
              summarize(winter_filings_20_25 = n())) |> 
  # Winter Executions 20-25
  left_join(executions_20_25_nycc |> 
              filter(!is.na(CounDist) & winter_season == TRUE) |> 
              group_by(CounDist) |> 
              summarize(winter_executions_20_25 = n())) |> 
  # Filings 25
  left_join(filings_20_25_nycc |> 
              filter(!is.na(CounDist) & year(fileddate) == 2025) |> 
              group_by(CounDist) |> 
              summarize(filings_25 = n())) |>
  # Executions 25
  left_join(executions_20_25_nycc |> 
              filter(!is.na(CounDist) & year(executiondate) == 2025) |> 
              group_by(CounDist) |> 
              summarize(executions_25 = n())) |> 
  # Winter filings 25
  left_join(filings_20_25_nycc |> 
              filter(!is.na(CounDist) & winter_season == TRUE & year(fileddate) == 2025) |> 
              group_by(CounDist) |> 
              summarize(winter_filings_25 = n())) |>
  #Winter Executions 25
  left_join(executions_20_25_nycc |> 
              filter(!is.na(CounDist) & winter_season == TRUE & year(executiondate) == 2025) |> 
              group_by(CounDist) |> 
              summarize(winter_executions_25 = n())) |> 
  st_drop_geometry()

write_xlsx(list("NYS Assembly" = nysad_compiled,
                "NYS Senate" = nyss_compiled,
                "NYC Council" = nyc_council_compiled),
          path = "NYS_Evictions_2020_2025.xlsx")

```

