---
title: "RTC Fact Sheet: 2025 State of Evictions"
format: 
  html:
    code-fold: true
---

```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false

# Insure packages are installed by checking required packages against installed packages
reqd_pkgs <- c("flextable",
               "sf",
               "tidyverse",
               "mapgl",
               "htmltools",
               "officer",
               "forcats",
               "DBI",
               "RPostgres",
               "socratadata",
               "lubridate",
               "writexl")

install.packages(
  setdiff(
    reqd_pkgs,
    rownames(installed.packages())
  ),
  repos = "https://cloud.r-project.org"
)

lapply(reqd_pkgs, library, character.only = TRUE) |> 
  invisible()

# Ensure that the Quarto render and RStudio read the same environment vars
readRenviron(".Renviron")

con <- dbConnect(
  RPostgres::Postgres(),
  dbname = Sys.getenv("PG_DBNAME"),
  host     = Sys.getenv("PG_HOST"),
  port     = Sys.getenv("PG_PORT"),
  user     = Sys.getenv("PG_USER"),
  password = Sys.getenv("PG_PASSWORD")
  )

set_flextable_defaults(
  font.family  = "Authentic Sans",
  font.size    = 10,        # ~13px (flextable uses points)
  border.color = "#0b0b0b",
  background.color = "transparent",
  padding.top    = 8,
  padding.bottom = 4,
  padding.left   = 0,
  padding.right  = 16
)

# Wraps a data frame in a CIL-styled flextable ready to embed in the page.
# Columns size to their content by default. `widths` overrides that: one value
# per column, or named values setting just those columns. `max_width` caps the
# fitted widths, never the ones `widths` names. `full_height = TRUE` drops the
# 400px cap so the whole table renders inline without a vertical scrollbar.
# `unit = "%"` reads `widths` as shares of a full-width table instead of
# absolute sizes, which is how two tables are given matching columns.
# `fill` is a named vector of cell text -> background colour, letting a
# category be scanned down a column by colour as well as read.
# Usage: data |> cil_table(widths = c("Purpose of Funds" = 4))
#        data |> cil_table(widths = c(39, 61), unit = "%")
#        data |> cil_table(fill = c("Heavy Rain" = "#d1ddee"))
cil_table <- function(data, widths = NULL, heights = NULL,
                      max_width = NULL, unit = "in", full_width = TRUE,
                      full_height = FALSE, fill = NULL) {
  ft <- flextable(data)
  pct <- identical(unit, "%")
  # Percentages never reach flextable, so they don't take the fitted path below.
  fitted <- !pct && !(is.null(widths) && is.null(max_width))

  if (pct) {
    if (is.null(widths)) {
      stop("`unit = \"%\"` sets column shares, so it needs `widths`.", call. = FALSE)
    }
    if (!is.null(max_width)) {
      stop("`max_width` caps fitted widths in absolute units; `unit = \"%\"` ",
           "sets shares outright. Use one or the other.", call. = FALSE)
    }
    if (!full_width) {
      stop("Percentage widths are shares of a full-width table. Leave ",
           "`full_width = TRUE`.", call. = FALSE)
    }
    if (!is.null(heights)) {
      # `unit` covers heights too, and a row can't be a percentage of anything.
      stop("`heights` needs an absolute unit, which `unit = \"%\"` overrides. ",
           "Set column shares and row heights in separate calls.", call. = FALSE)
    }
    if (length(widths) != ncol(data)) {
      # Named subsets work for absolute widths because dim_pretty() supplies a
      # baseline for the columns left out. Shares have no such baseline.
      stop("`widths` has ", length(widths), " value(s) but `data` has ",
           ncol(data), " column(s). Percentage widths need one per column.",
           call. = FALSE)
    }
  }

  if (fitted && full_width) {
    stop("`full_width` stretches columns to the container; `widths`/`max_width` ",
         "pin them. Use one or the other.", call. = FALSE)
  }

  if (!fitted) {
    # Browser-side content fitting, like Excel's Autofit. width = 0 leaves the
    # table unconstrained; width = 1 would stretch it to 100% regardless of text.
    ft <- set_table_properties(ft, if (full_width) 1 else 0, layout = "autofit")
  } else {
    # Fitted widths are the baseline, capped by max_width, then overridden by
    # anything `widths` names. "autofit" discards widths, so this must be fixed.
    fit <- unlist(dim_pretty(ft)$widths)
    if (!is.null(max_width)) fit <- pmin(fit, max_width)
    if (is.null(widths)) {
      widths <- fit
    } else if (!is.null(names(widths)) && all(nzchar(names(widths)))) {
      unknown <- setdiff(names(widths), names(data))
      if (length(unknown)) {
        stop("`widths` names no such column: ", paste(unknown, collapse = ", "),
             call. = FALSE)
      }
      fit[match(names(widths), names(data))] <- widths
      widths <- fit
    } else if (length(widths) != ncol(data)) {
      # Otherwise they land on the first columns, misaligning every one.
      stop("`widths` has ", length(widths), " value(s) but `data` has ",
           ncol(data), " column(s). Give one width per column, or name them.",
           call. = FALSE)
    }
    ft <- set_table_properties(ft, layout = "fixed")
    ft <- width(ft, j = seq_along(widths), width = widths, unit = unit)
  }

  if (!is.null(heights)) {
    # Heights are only emitted under the "exact" rule; "auto" ignores them.
    ft <- height(ft, height = heights, part = "body", unit = unit)
    ft <- hrule(ft, rule = "exact", part = "body")
  }

  if (!is.null(fill)) {
    unknown <- setdiff(names(fill), unlist(lapply(data, as.character)))
    if (length(unknown)) {
      stop("`fill` names a value no cell holds: ", paste(unknown, collapse = ", "),
           call. = FALSE)
    }
    # bg() hands the function one column at a time, so the lookup is on the cell
    # text itself. Anything unnamed -- ids, counts -- keeps the table's own
    # background rather than picking up a colour by accident.
    ft <- bg(ft, part = "body", bg = function(x) {
      matched <- fill[as.character(x)]
      ifelse(is.na(matched), "transparent", matched)
    })
  }

  # flextable's width() speaks only in/cm/mm, and autofit emits no column widths
  # at all, so shares are carried by markup instead: a <colgroup> the browser
  # obeys once .scroll-output--pct switches the table to `table-layout: fixed`.
  # Empty otherwise, leaving the sub() below a no-op.
  colgroup <- if (pct) {
    paste0("<colgroup>",
           paste0(sprintf('<col style="width:%s%%;">', widths), collapse = ""),
           "</colgroup>")
  } else ""

  ft |>
    border_remove() |>
    bold(bold = TRUE, part = "header") |>
    hline(border = fp_border(color = "#0b0b0b", width = 1.5),  part = "header") |>
    hline(border = fp_border(color = "#969696", width = 0.75), part = "body") |>
    # hline() mirrors each rule onto the next row's top edge as well as the
    # current row's bottom. Under the `border-collapse: separate` these tables
    # need for sticky headers, the two edges stack instead of merging, so every
    # interior rule renders at double weight -- only the final one, with no row
    # below it, looks right. Clearing the tops leaves a single 0.75pt line.
    border(border.top = fp_border(width = 0), part = "body") |>
    padding(
      padding.top = 0, padding.bottom = 4,
      padding.left = 0, padding.right = 8,
      part = "header"
    ) |>
    padding(
      padding.left = 2, padding.right = 8,
      part = "body"
    ) |> 
    # Both steps are needed to keep the table out of a shadow DOM, which no page
    # CSS can reach: ft.shadow drops the static host, and the rename hides the
    # wrapper from tabwid.js, which shadow-roots every .tabwid on DOMContentLoaded.
    htmltools_value(ft.shadow = FALSE) |>
    as.character() |>
    sub(pattern = 'class="tabwid"', replacement = 'class="cil-tabwid"', fixed = TRUE) |>
    sub(pattern = "(<table[^>]*>)", replacement = paste0("\\1", colgroup)) |>
    HTML() |>
    div(class = paste(
      c("scroll-output",
        if (fitted) "scroll-output--fit",
        if (pct) "scroll-output--pct",
        if (full_height) "scroll-output--full"),
      collapse = " "
    ))
}
```


The following analyses were performed in support of [Right to Counsel's](https://www.righttocounselnyc.org/) annual State of Evictions fact sheet. 

```{r, cache=TRUE, message=FALSE}
# First, pull data. Then, work locally in R.

### Evictions db SQL queries ###

# Eviction Filings, 2020 - 2025, all
filings_20_25 <- dbGetQuery(con, "
  SELECT DISTINCT(i.indexnumberid), fileddate, court
  FROM oca_index i
  WHERE i.fileddate >= MAKE_DATE(2020,1,1) 
      AND i.fileddate <= MAKE_DATE(2025,12,31)
      AND i.classification IN ('Holdover', 'Non-Payment')
      AND (i.propertytype = 'Residential' OR i.propertytype IS NULL);
")

# Eviction Filings, 2020 - 2025, with location
filings_20_25_sf <- st_read(con, query = "
  SELECT DISTINCT(i.indexnumberid), fileddate, court,
    ST_Transform(a.geom, 2263) AS geom
  FROM oca_index i
  JOIN oca_addresses a ON i.indexnumberid = a.indexnumberid
  WHERE i.fileddate >= MAKE_DATE(2020,1,1) 
      AND i.fileddate <= MAKE_DATE(2025,12,31)
      AND i.classification IN ('Holdover', 'Non-Payment')
      AND (i.propertytype = 'Residential' OR i.propertytype IS NULL)
      AND a.geom IS NOT NULL;
")

# Executed Evictions, 2020 - 2025, all
executions_20_25 <- dbGetQuery(con, "
  SELECT DISTINCT(i.indexnumberid), w.executiondate, i.court
  FROM oca_index i
  JOIN oca_warrants w ON i.indexnumberid = w.indexnumberid
  WHERE w.executiondate >= MAKE_DATE(2020,1,1) 
    AND w.executiondate <= MAKE_DATE(2025,12,31)
    AND i.classification IN ('Holdover', 'Non-Payment')
    AND (i.propertytype = 'Residential' OR i.propertytype IS NULL)
    AND w.executiondate IS NOT NULL;
")

# Executed Evictions, 2020 - 2025, with location
executions_20_25_sf <- st_read(con, query = "
  WITH executed_evictions AS (
  	SELECT DISTINCT(i.indexnumberid), w.executiondate, i.court
  	FROM oca_index i
  	JOIN oca_warrants w ON i.indexnumberid = w.indexnumberid
  	WHERE w.executiondate >= MAKE_DATE(2020,1,1) 
  	  AND w.executiondate <= MAKE_DATE(2025,12,31)
  	  AND i.classification IN ('Holdover', 'Non-Payment')
  	  AND (i.propertytype = 'Residential' OR i.propertytype IS NULL)
  	  AND w.executiondate IS NOT NULL
  )
  SELECT e.*, ST_TRANSFORM(a.geom, 2263)
  FROM executed_evictions e
  JOIN oca_addresses a on e.indexnumberid = a.indexnumberid
  WHERE a.geom IS NOT NULL
")

### Boundary Files ###
# NYC Boundary via borough boundary union
nyc <- read_sf("https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_Borough_Boundary/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson") |> 
  st_transform(2263) |> 
  st_make_valid() |> 
  st_union()

# NYS Assembly
nysad <- read_sf("https://services6.arcgis.com/EbVsqZ18sv1kVJ3k/ArcGIS/rest/services/NYS_Assembly_Districts/FeatureServer/0/query?where=0%3D0&outFields=*&f=geojson") |> 
  st_transform(2263) |> 
  st_make_valid()

# NYS Senate
nyss <- read_sf("https://services6.arcgis.com/EbVsqZ18sv1kVJ3k/ArcGIS/rest/services/NYS_Senate_Districts/FeatureServer/0/query?where=0%3D0&outFields=*&f=geojson") |> 
  st_transform(2263) |> 
  st_make_valid()

# NYC Council
nyc_council <- read_sf("https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_City_Council_Districts/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson") |> 
  st_transform(2263) |> 
  st_make_valid()

# NYC Council members
nyc_council_members <- soc_read(
  "https://data.cityofnewyork.us/City-Government/City-Council-Members-1999-to-Present-/uvw5-9znb.json"
) |> 
  # Filter to current term
  filter(Sys.Date() %within% interval(start = term_start, end = term_end))

# Join CMs to spatial file
nyc_council <- nyc_council |> 
  left_join(nyc_council_members |> 
              select(name, district),
            by = join_by(CounDist == district)) |> 
  # Designate Vacant seats
  mutate(name = replace_na(name, "Vacant"))

# NYC Boundaries via Borough union
bb <- read_sf("https://services5.arcgis.com/GfwWNkhOj9bNBqoJ/arcgis/rest/services/NYC_Borough_Boundary/FeatureServer/0/query?where=1=1&outFields=*&outSR=4326&f=pgeojson") |> 
  st_transform(2263) |> 
  st_make_valid() |> 
  st_union()

# NYS Counties
nys_counties <- read_sf("https://services6.arcgis.com/EbVsqZ18sv1kVJ3k/ArcGIS/rest/services/NYS_Civil_Boundaries/FeatureServer/2/query?where=0%3D0&outFields=*&f=geojson") |> 
  st_transform(2263)

# Separate out upstate / downstate
nysad_downstate <- st_filter(nysad, bb, .predicate = st_intersects)

nysad_upstate <- st_filter(nysad, bb, .predicate = st_disjoint)

nyss_downstate <- st_filter(nyss, bb, .predicate = st_intersects)
  
nyss_upstate <- st_filter(nyss, bb, .predicate = st_disjoint)
```


## Cumulative Eviction Filings

### Number of eviction filings from January 1, 2020 - December 31, 2025

In New York State, there were **`r scales::comma(nrow(filings_20_25))`** eviction-oriented (holdover and non-payment) filings from January 1, 2020 to December 31, 2025.

Please note that this analysis **excludes commercial evictions within New York City.** Unfortunately, outside of NYC, excluding commercial from residential eviction filings and executions is not possible given the current data.

## Annual eviction filings
```{r}
eviction_count_annual_df <- filings_20_25 |> 
  group_by(filed_year = year(fileddate)) |> 
  summarize(count = n())
```

The number of eviction-related (holdover and non-payment) eviction filings per year is:

```{r}
eviction_count_annual_table <- eviction_count_annual_df |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    filed_year = "Year",
    count = "Eviction Filings"
  ) |> 
  colformat_num(j=1,
                big.mark = "") |> 
  align(j=1,
        align = "left",
        part = "all")

eviction_count_annual_table
```

## Which areas are hardest hit?

### By court, January 1, 2020 - December 31, 2025:
```{r}
eviction_count_courts_df <- filings_20_25 |> 
  group_by(court) |> 
  summarize(count = n()) |> 
  arrange(desc(count))

eviction_count_courts_table <- eviction_count_courts_df |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    court = "Court",
    count = "Eviction Filings"
  )

eviction_count_courts_table
```

### By court, January 1, 2025 - December 31, 2025:
```{r}
eviction_count_courts_25_df <- filings_20_25 |> 
  filter(year(fileddate) == 2025) |> 
  group_by(court) |> 
  summarize(count = n()) |> 
  arrange(desc(count))

eviction_count_courts_25_table <- eviction_count_courts_25_df |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    court = "Court",
    count = "Eviction Filings"
  )

eviction_count_courts_25_table
```

### By NYS Assembly District: January 1, 2020 - December 31, 2025

*In the following filing counts by area, note that only `r scales::comma(nrow(filings_20_25_sf))` eviction filings (`r round((nrow(filings_20_25_sf)/nrow(filings_20_25)*100),2)`%) have a location*

#### Upstate Assembly Districts (Not intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nysad_upstate_5year <- filings_20_25_sf |> 
  st_join(nysad_upstate |> 
              select(District, Name),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(District, Name) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nysad_upstate_5year_table <- nysad_upstate_5year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    Name = "Representative"
  )

nysad_upstate_5year_table
```

#### Downstate Assembly Districts (Intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nysad_downstate_5year <- filings_20_25_sf |> 
  st_join(nysad_downstate |> 
              select(District, Name),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(District, Name) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nysad_downstate_5year_table <- nysad_downstate_5year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all")

nysad_downstate_5year_table
```

### By NYS Assembly District: January 1, 2025 - December 31, 2025

#### Upstate Assembly Districts (Not intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nysad_upstate_1year <- filings_20_25_sf |> 
  filter(year(fileddate) == 2025) |> 
  st_join(nysad_upstate |> 
              select(District, Name),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(District, Name) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nysad_upstate_1year_table <- nysad_upstate_1year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all")

nysad_upstate_1year_table
```

#### Downstate Assembly Districts (Intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nysad_downstate_1year <- filings_20_25_sf |> 
  filter(year(fileddate) == 2025) |>  
  st_join(nysad_downstate |> 
              select(District, Name),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(District, Name) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`)) 

nysad_downstate_1year_table <- nysad_downstate_1year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all")

nysad_downstate_1year_table
```

### By NYS Senate District, January 1, 2020 - December 31, 2025

#### Upstate Senate Districts (Not intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nyss_upstate_5year <- filings_20_25_sf |> 
  st_join(nyss_upstate |> 
              select(DISTRICT, NAME),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(DISTRICT, NAME) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`)) 

nyss_upstate_5year_table <- nyss_upstate_5year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    DISTRICT = "District",
    NAME = "Representative"
  )

nyss_upstate_5year_table
```

#### Downstate Senate Districts (Intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nyss_downstate_5year <- filings_20_25_sf |> 
  st_join(nyss_downstate |> 
              select(DISTRICT, NAME),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(DISTRICT, NAME) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`)) 

nyss_downstate_5year_table <- nyss_downstate_5year |>
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    DISTRICT = "District",
    NAME = "Representative"
  )

nyss_downstate_5year_table
```

### By NYS Senate District, January 1, 2025 - December 31, 2025
#### Upstate Senate Districts (Not intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nyss_upstate_1year <- filings_20_25_sf |> 
  filter(year(fileddate) == 2025) |> 
  st_join(nyss_upstate |> 
              select(DISTRICT, NAME),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(DISTRICT, NAME) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nyss_upstate_1year_table <- nyss_upstate_1year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    DISTRICT = "District",
    NAME = "Representative"
  )

nyss_upstate_1year_table
```

#### Downstate Senate Districts (Intersecting with NYC Borough boundaries)
```{r, cache=FALSE}
nyss_downstate_1year <- filings_20_25_sf |> 
  filter(year(fileddate) == 2025) |> 
  st_join(nyss_downstate |> 
              select(DISTRICT, NAME),
            left = FALSE) |> 
  st_drop_geometry() |> 
  group_by(DISTRICT, NAME) |> 
  summarize(`Eviction Filings` = n()) |> 
  arrange(desc(`Eviction Filings`))

nyss_downstate_1year_table <- nyss_downstate_1year |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    DISTRICT = "District",
    NAME = "Representative"
  )

nyss_downstate_1year_table
```

### Executed Evictions in 2025

Note: Of the `r scales::comma(nrow(executions_20_25 |> filter(year(executiondate)==2025)))` evictions executed statewide in 2025, only `r scales::comma(nrow(executions_20_25_sf |> filter(year(executiondate)==2025)))` have a location associated with them (`r round((nrow(executions_20_25_sf |> filter(year(executiondate)==2025))/nrow(executions_20_25 |> filter(year(executiondate)==2025))*100),2)`%)

``` {r, cache=FALSE}
executed_evictions_25_byCty <- executions_20_25_sf |> 
  filter(year(executiondate) == 2025) |> 
  st_join(nys_counties |> 
            select(NAME)) |> 
  st_drop_geometry() |> 
  group_by(NAME) |> 
  summarize(`Executed Evictions` = n()) |> 
  arrange(desc(`Executed Evictions`))

executed_evictions_25_byCty_table <- executed_evictions_25_byCty |> 
  head(10) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  align(j=1,
        align = "left",
        part = "all") |> 
  set_header_labels(
    NAME = "County"
  )

executed_evictions_25_byCty_table
```


### Evictions During the Winter Months (November 1 - April 15)

#### Eviction filings

```{r}
filings_20_25 <- filings_20_25 |> 
  mutate(winter_season = case_when(month(fileddate) >= 11 ~ TRUE,
                                   month(fileddate) < 4 ~ TRUE,
                                   month(fileddate) == 4 & day(fileddate) <= 15 ~ TRUE,
                                   .default = FALSE
                                   ))

winter_filings_table <- filings_20_25 |> 
  group_by(filed_year = year(fileddate)) |> 
  summarize(all_filings = n(),
            winter_filings = sum(winter_season)) |> 
  mutate(winter_filing_pct = round((winter_filings/all_filings*100), 1)) |> 
  flextable() |> 
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    filed_year = "Year",
    all_filings = "Annual Filings",
    winter_filings = "Winter Filings",
    winter_filing_pct = "Winter Filing Percentage (%)"
  ) |> 
  colformat_num(j=1,
                big.mark = "") |> 
  colformat_int(j=2:3,
                big.mark = ",") |> 
  align(j=1,
      align = "left",
      part = "all")

winter_filings_table
```

#### Executed evictions

```{r}
executions_20_25 <- executions_20_25 |> 
  mutate(winter_season = case_when(month(executiondate) >= 11 ~ TRUE,
                                   month(executiondate) < 4 ~ TRUE,
                                   month(executiondate) == 4 & day(executiondate) <= 15 ~ TRUE,
                                   .default = FALSE
                                   ))

winter_executions_table <- executions_20_25 |> 
  group_by(executed_year = year(executiondate)) |>
  summarize(all_executions = n(),
            winter_executions = sum(winter_season)) |> 
  mutate(winter_execution_pct = round(winter_executions/all_executions*100,1)) |> 
  flextable() |>
  set_table_properties(width = 1, layout = "autofit") |> 
  set_header_labels(
    executed_year = "Year",
    all_executions = "Annual Executions",
    winter_executions = "Winter Executions",
    winter_execution_pct = "Winter Execution Percentage (%)"
  ) |> 
  colformat_num(j=1,
                big.mark = "") |> 
  colformat_int(j=2:3,
                big.mark = ",") |> 
  align(j=1,
      align = "left",
      part = "all")
  
winter_executions_table 
```

Compiling filings, executions, winter filings, and winter executions for the two political districts plus NYC Council districts:
```{r}
# Compile filings, executions, winter filings, winter executions for all political districts
filings_20_25_sf <- filings_20_25_sf |> 
  mutate(winter_season = case_when(month(fileddate) >= 11 ~ TRUE,
                                   month(fileddate) < 4 ~ TRUE,
                                   month(fileddate) == 4 & day(fileddate) <= 15 ~ TRUE,
                                   .default = FALSE
                                   ))

executions_20_25_sf <- executions_20_25_sf |> 
  mutate(winter_season = case_when(month(executiondate) >= 11 ~ TRUE,
                                   month(executiondate) < 4 ~ TRUE,
                                   month(executiondate) == 4 & day(executiondate) <= 15 ~ TRUE,
                                   .default = FALSE
                                   ))

### Assembly Districts ###
nysad_compiled <- nysad |> 
  select(District, Name, Party) |> 
  mutate(Region = case_when(District %in% nysad_upstate$District ~ "Upstate",
                            District %in% nysad_downstate$District ~ "Downstate",
                            .default = NA))

executions_20_25_nysad <- executions_20_25_sf |> 
  st_join(nysad_compiled |> select(District))



nysad_compiled <- nysad_compiled |> 
  left_join(
    bind_rows(
      nysad_downstate_5year,
      nysad_upstate_5year
    ) |> 
      rename("filings_20_25" = "Eviction Filings") |> 
      select(!Name),
    by = join_by(District)
  ) |> 
  left_join(
    bind_rows(
      nysad_downstate_1year,
      nysad_upstate_1year
    ) |> 
      rename("filings_25" = "Eviction Filings") |> 
      select(!Name),
    by = join_by(District)
  ) |> 
  left_join(executions_20_25_nysad |> 
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(executions_20_25 = n())) |> 
  left_join(executions_20_25_nysad |> 
              filter(year(executiondate) == 2025) |> 
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(executions_25 = n())) |> 
  left_join(filings_20_25_sf |> 
              filter(winter_season == TRUE) |> 
              st_join(nysad_compiled |> select(District)) |>
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(winter_filings_20_25 = n())) |> 
  left_join(filings_20_25_sf |> 
              filter(winter_season == TRUE & year(fileddate) == 2025) |> 
              st_join(nysad_compiled |> select(District)) |>
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(winter_filings_25 = n())) |> 
  left_join(executions_20_25_nysad |> 
              filter(winter_season == TRUE) |> 
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(winter_executions_20_25 = n())) |> 
  left_join(executions_20_25_nysad |> 
              filter(winter_season == TRUE & year(executiondate) == 2025) |> 
              st_drop_geometry() |> 
              group_by(District) |> 
              summarize(winter_executions_25 = n())) |> 
  st_drop_geometry()

# Rearrange
nysad_compiled <- nysad_compiled |> 
  relocate(contains("20_25"),
           .after = Region)

### Senate Districts ###
nyss_compiled <- nyss |> 
  select(DISTRICT, NAME, Party) |> 
  mutate(Region = case_when(DISTRICT %in% nyss_upstate$DISTRICT ~ "Upstate",
                            DISTRICT %in% nyss_downstate$DISTRICT ~ "Downstate",
                            .default = NA))

executions_20_25_nyss <- executions_20_25_sf |> 
  st_join(nyss_compiled |> select(DISTRICT))

nyss_compiled <- nyss_compiled |> 
  left_join(
    bind_rows(
      nyss_upstate_5year,
      nyss_downstate_5year
    ) |> 
      rename("filings_20_25" = "Eviction Filings") |> 
      select(!NAME),
    by = join_by(DISTRICT)
  ) |> 
  left_join(
    bind_rows(
      nyss_downstate_1year,
      nyss_upstate_1year
    ) |> 
      rename("filings_25" = "Eviction Filings") |> 
      select(!NAME),
    by = join_by(DISTRICT)
  ) |> 
  left_join(executions_20_25_nyss |> 
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(executions_20_25 = n())) |> 
  left_join(executions_20_25_nyss |> 
              filter(year(executiondate) == 2025) |> 
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(executions_25 = n())) |> 
  left_join(filings_20_25_sf |> 
              filter(winter_season == TRUE) |> 
              st_join(nyss_compiled |> select(DISTRICT)) |>
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(winter_filings_20_25 = n())) |> 
  left_join(filings_20_25_sf |> 
              filter(winter_season == TRUE & year(fileddate) == 2025) |> 
              st_join(nyss_compiled |> select(DISTRICT)) |>
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(winter_filings_25 = n())) |> 
  left_join(executions_20_25_nyss |> 
              filter(winter_season == TRUE) |> 
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(winter_executions_20_25 = n())) |> 
  left_join(executions_20_25_nyss |> 
              filter(winter_season == TRUE & year(executiondate) == 2025) |> 
              st_drop_geometry() |> 
              group_by(DISTRICT) |> 
              summarize(winter_executions_25 = n())) |> 
  st_drop_geometry()
  
# Rearrange
nyss_compiled <- nyss_compiled |> 
  relocate(contains("20_25"),
           .after = Region)


### NYC Council ###
nyc_council <- nyc_council |> 
  select(CounDist, name) |> 
  arrange(CounDist)

filings_20_25_nycc <- filings_20_25_sf |> 
  st_join(nyc_council |> select(CounDist)) |> 
  st_drop_geometry()

executions_20_25_nycc <- executions_20_25_sf |> 
  st_join(nyc_council |> select(CounDist)) |> 
  st_drop_geometry()

nyc_council_compiled <- nyc_council |> 
  # Filings 20-25
  left_join(filings_20_25_nycc |> 
              filter(!is.na(CounDist)) |> 
              group_by(CounDist) |> 
              summarize(filings_20_25 = n())) |> 
  # Executions 20-25
  left_join(executions_20_25_nycc |> 
              filter(!is.na(CounDist)) |> 
              group_by(CounDist) |> 
              summarize(executions_20_25 = n())) |> 
  # Winter filings 20-25
  left_join(filings_20_25_nycc |> 
              filter(!is.na(CounDist) & winter_season == TRUE) |> 
              group_by(CounDist) |> 
              summarize(winter_filings_20_25 = n())) |> 
  # Winter Executions 20-25
  left_join(executions_20_25_nycc |> 
              filter(!is.na(CounDist) & winter_season == TRUE) |> 
              group_by(CounDist) |> 
              summarize(winter_executions_20_25 = n())) |> 
  # Filings 25
  left_join(filings_20_25_nycc |> 
              filter(!is.na(CounDist) & year(fileddate) == 2025) |> 
              group_by(CounDist) |> 
              summarize(filings_25 = n())) |>
  # Executions 25
  left_join(executions_20_25_nycc |> 
              filter(!is.na(CounDist) & year(executiondate) == 2025) |> 
              group_by(CounDist) |> 
              summarize(executions_25 = n())) |> 
  # Winter filings 25
  left_join(filings_20_25_nycc |> 
              filter(!is.na(CounDist) & winter_season == TRUE & year(fileddate) == 2025) |> 
              group_by(CounDist) |> 
              summarize(winter_filings_25 = n())) |>
  #Winter Executions 25
  left_join(executions_20_25_nycc |> 
              filter(!is.na(CounDist) & winter_season == TRUE & year(executiondate) == 2025) |> 
              group_by(CounDist) |> 
              summarize(winter_executions_25 = n())) |> 
  st_drop_geometry()

write_xlsx(list("NYS Assembly" = nysad_compiled,
                "NYS Senate" = nyss_compiled,
                "NYC Council" = nyc_council_compiled),
          path = "NYS_Evictions_2020_2025.xlsx")

```

### Compiled data by district

```{r}
# Display labels for the count columns shared by all three compiled tables
compiled_labels <- c(
  "Filings 2020-25"            = "filings_20_25",
  "Executions 2020-25"         = "executions_20_25",
  "Winter Filings 2020-25"     = "winter_filings_20_25",
  "Winter Executions 2020-25"  = "winter_executions_20_25",
  "Filings 2025"               = "filings_25",
  "Executions 2025"            = "executions_25",
  "Winter Filings 2025"        = "winter_filings_25",
  "Winter Executions 2025"     = "winter_executions_25"
)

# Counts are doubles after the joins; as integers they print without decimals
compiled_counts <- function(data) {
  data |>
    mutate(across(any_of(unname(compiled_labels)), as.integer)) |>
    rename(any_of(compiled_labels))
}
```

#### NYS Assembly Districts

```{r}
nysad_compiled |>
  arrange(District) |>
  compiled_counts() |>
  rename(Representative = Name) |>
  cil_table()
```

#### NYS Senate Districts

```{r}
nyss_compiled |>
  arrange(DISTRICT) |>
  compiled_counts() |>
  rename(District = DISTRICT,
         Representative = NAME) |>
  cil_table()
```

#### NYC Council Districts

```{r}
nyc_council_compiled |>
  arrange(CounDist) |>
  compiled_counts() |>
  rename(District = CounDist,
         `Council Member` = name) |>
  cil_table()
```

*Any data provided herein does not constitute an official record of the New York State Unified Court
System, which does not represent or warrant the
accuracy thereof. The opinions, findings, and
conclusions expressed in this publication are those of
the authors and not those of the New York State
Unified Court System, which assumes no liability for its
contents or use thereof.* 
